---
title: "20231030_AssignmentB1"
author: "Torin"
date: '2023-10-31'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

_We will begin by loading the packages required for this code_
```{r}
library(tidyverse)
library(testthat)
library(gapminder)
```

# Creating a Function

_The following code will create a function that calculates the standard deviation of a vector of numerics representing a sample_
```{r}
#' Title: stdev()
#'
#' @description This function calculates the standard deviation of a vector of numbers, automatically removing any missing values. The input must be numeric, or the function will throw an error.
#' 
#' @param w a numeric vector, named 'w' in keeping with conventions for vectors
#' 
#' @return The function returns the standard deviation of the vector. The mean and length of the vector are also printed as the function runs, to double-check that the values used for calculating the standard deviation are correct.

stdev <- function(w) {
  stopifnot(is.numeric(w))
  w = w[!is.na(w)]
  mu = mean(w)
  print(paste("Mean:", mu))
  n = length(w)
  print(paste("Length:", n))
  sum_sq = 0
  for (i in 1:length(w)) {
    sum_sq = sum_sq + (w[i] - mu)^2
  }
  sd_sq = sum_sq/(n - 1)
  stdeviation = sqrt(sd_sq)
  print("Standard deviation:")
  return(stdeviation)
}
```

# Examples

_The following code will demonstrate the use of the newly created function `stdev()` using three examples_
```{r}
#Example 1: calculate the standard deviation of a vector
vec <- c(2,3,4,5,10,12)
stdev(w = vec)

#Example 2: calculate the standard deviation of a vector with missing values
vec1 <- c(2,3,4,5,NA,NA,10,12)
stdev(w = vec1)

#Example 3: Calculate the standard deviation of a numeric column in the gapminder dataset
stdev(gapminder$lifeExp)
```
# Testing

_The following code will formally test that the function returns the expected output in three distinct scenarios_
```{r}
# Test that the function output is the same as the built-in dplyr function sd()
test_that("vec", {expect_equal(sd(vec), stdev(vec))})

#Test that the function returns an error if w is not numeric
vec2 <- c("apple", "banana", "orange")
test_that("vec", {expect_error(stdev(vec2))})

#Test that the function returns the same output regardless of missing values in the vector
test_that("vec", {expect_equal(stdev(vec), stdev(vec1))})

```